Experiments with rtrees using `CCFs_withlands` data

### 2020-09-25 journals/01-hamby_data_and_models/01-hamby_data_and_models.Rmd
everything is a mess right now, but pushing this so it doesn't get lost

If we include symmetric comparisons, each set of test bullets should result in a dataset with
  $$(35 \mbox{ bullets} \times 6 \mbox{ lands})^2=44100 \mbox{ rows},$$
where a row would contain information on a pair of lands. If we do not include the symmetric comparisons, then the dataset should have
  $$\frac{(44100 \mbox{ rows} - (35 \mbox{ bullets} \times 6))}{2} + (35 \mbox{ bullets} \times 6) = 22155 \mbox{ rows}.$$
However, when I looked at the dimension of the datasets, neither of these seem to be the case. See the R code and output below. Note that `hamby173` is currently incorrectly labelled as `hamby44`. Both test sets have less than but close to 22,155 rows. This suggests that these do not include symmetric comparisons. When I checked with Heike, she confirmed that this is the case. This table also shows that there are comparisons across `hamby173` and `hamby252`. These missing observations will be explored more in the next section.

```{r}
# Summary of the number of observations in the Hamby173and252 datase
table(hamby173and252_raw$study1, hamby173and252_raw$study2)
```

@@

The plot below considers the number of observations within a barrel and bullet comparison from all known cases in the Hamby 173 and 252 data. We can see that the observations on the lower diagonals are missing in all cases which confirms that the symmetric comparisons were not included in the data. Additionally, a handful of cases have less than 36 observations. For the comparisons within the Hamby 173 or Hamby 252 study, the cells on the diagonals are less than 36, because none of the repeats from the symmetric comparisons of lands are included. The cells above the diagonal with less than 36 observations are missing some observations due to tank rash. For the comparisons across studies, the cases with less than expected are also due to tank rash. For some reason, the comparisons between bullets 1 from Hamby 173 and Hamby 252, the cells are being colored grey even though they have 36 observations. I am not sure why this is...

@@

```{r}
# Determine the letters associated with the unknown bullets
letters <- levels(hamby173and252_raw$barrel1)[11:length(levels(hamby173and252_raw$barrel1))]

# Cleaning the testing data
hamby173and252_train_cleaning <- hamby173and252_raw %>%
  mutate(study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
         study2 = fct_recode(study2, "Hamby173" = "Hamby44"),
         bullet1 = factor(ifelse(barrel1 %in% letters,
                                 as.character(barrel1),
                                 as.character(bullet1))),
         barrel1 = factor(ifelse(barrel1 %in% letters, 
                                 as.character("Unknown"), 
                                 as.character(barrel1))),
         bullet2 = factor(ifelse(barrel2 %in% letters,
                                 as.character(barrel2),
                                 as.character(bullet2))),
         barrel2 = factor(ifelse(barrel2 %in% letters, 
                                 as.character("Unknown"), 
                                 as.character(barrel2))),
         land1 = factor(land1),
         land2 = factor(land2),
         rfscore = predict(bulletxtrctr::rtrees, hamby173and252_raw %>%
                             select(rf_features), 
                           type = "prob")[,2]) %>%
  rename(samesource = match) %>%
  select(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2,
         rf_features, samesource, rfscore)
```

@@

```{r}
# Remove the comparisons involving the four lands with tank rash
hamby173and252_train <- hamby173and252_train_cleaning %>%
  mutate(bbl1 = barrel1:bullet1:land1,
         bbl2 = barrel2:bullet2:land2) %>%
  filter(!(bbl1 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4") | 
           bbl2 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4"))) %>%
  select(-bbl1, -bbl2)
```

@@

```{r}
# Create dataset for creating the paralle coordinate plot
par_data <- hamby173and252_train_cleaning %>%
  mutate(bbl1 = barrel1:bullet1:land1,
         bbl2 = barrel2:bullet2:land2) %>%
  mutate(category = factor(ifelse(!(bbl1 %in% c("6:2:1", "9:2:4", "Unknown:B:2",
                                                "Unknown:Q:4") | 
                                      bbl2 %in% c("6:2:1", "9:2:4", "Unknown:B:2",
                                                  "Unknown:Q:4")),
                           "keep", "remove"))) %>%
  arrange(category) %>%
  mutate(samesource_colors = ifelse(samesource == TRUE, rgb(1, 0, 0, alpha = 0.05), 
                                    rgb(0, 0, 0, alpha = 0.001)),
         samesource_colors2 = ifelse(samesource == TRUE, rgb(1, 0, 0, alpha = 0.001), 
                                    rgb(0, 0, 0, alpha = 0.01)),
         tankrash_colors = ifelse(category == "keep", rgb(0, 0, 1, alpha = 0.05),
                                  rgb(0, 0, 0, alpha = 0.01)))
```

The first plot shows the observations that were kept in black and the observations that were removed in blue. These observations do not stand out to me. It is not clear if these were the ones that were suppose to be removed or not. The second plot shows the observations from matches plotted in red, and the third plot shows the observations from nonmatches plotted in black.

@@

Below is a parallel coordinate plot of the training dataset. It is facetted by `samesource` and colored by `rfscore`. The features are sorted by the variable importance scores (descending from left to right) from the `rtrees` model. It is interesting to see that the matches with low rfscores have feature values that are similar to the nonmatches.

@@

Below is a parallel coordinate plot of the testing dataset. It is facetted by `samesource` and colored by `rfscore`. You can see a clear distinction between the trends of the feature values from the nonmatches and matches. It is interesting to consider the cases in set 1 that are matches but were predicted to be nonmatches by the model. These are the cases colored in grey, and their feature values are very different from the observations that were predicted to be matches by the model. The nonmatch predictions have values that are closer to the nonmatches.

@@

I have been thinking a lot about it is difficult to determine whether LIME is providing reasonable explanations, because we cannot know what the random forest is doing. That is the point of LIME. Heike suggested that we fit a logistic regression model to the training bullet data. We can then apply LIME to the logistic regression and determine if the explanations match what we would expect based on the coefficients of the logistic regression that we can interpret. She also suggested including interactions in the logistic regression model. This section includes code for fitting several logisitic regression models.

>>>>

If we include symmetric comparisons, each set of test bullets should result in a dataset with
  $$(35 \mbox{ bullets} \times 6 \mbox{ lands})^2=44100 \mbox{ rows},$$
where a row would contain information on a pair of lands. If we do not include the symmetric comparisons, then the dataset should have
  $$\frac{(44100 \mbox{ rows} - (35 \mbox{ bullets} \times 6))}{2} + (35 \mbox{ bullets} \times 6) = 22155 \mbox{ rows}.$$
However, when I looked at the dimension of the datasets, neither of these seem to be the case. See the R code and output below. Note that `hamby173` is currently incorrectly labeled as `hamby44`. Both test sets have less than but close to 22,155 rows. This suggests that these do not include symmetric comparisons. When I checked with Heike, she confirmed that this is the case. This table also shows that there are comparisons across `hamby173` and `hamby252`. These missing observations will be explored more in the next section.

```{r}
# Summary of the number of observations in the Hamby173and252 database
table(hamby173and252_raw$study1, hamby173and252_raw$study2)
```

@@

The plot below considers the number of observations within a barrel and bullet comparison from all known cases in the Hamby 173 and 252 data. We can see that the observations on the lower diagonals are missing in all cases which confirms that the symmetric comparisons were not included in the data. Additionally, a handful of cases have less than 36 observations. For the comparisons within the Hamby 173 or Hamby 252 study, the cells on the diagonals are less than 36, because none of the repeats from the symmetric comparisons of lands are included. The cells above the diagonal with less than 36 observations are missing some observations due to tank rash. For the comparisons across studies, the cases with less than expected are also due to tank rash. For some reason, the comparisons between bullets 1 from Hamby 173 and Hamby 252, the cells are being colored gray even though they have 36 observations. I am not sure why this is...

@@

```{r}
# Determine the letters associated with the unknown bullets
all_barrel_labels <-
  unique(c(
    as.character(hamby173and252_raw$barrel1),
    as.character(hamby173and252_raw$barrel2)
  ))
letters <- all_barrel_labels[!(all_barrel_labels %in% 1:10)]
letters

#letters <- levels(hamby173and252_raw$barrel1)[11:length(levels(hamby173and252_raw$barrel1))]

# Cleaning the testing data
hamby173and252_train_cleaning <- 
  hamby173and252_raw %>%
  mutate(study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
         study2 = fct_recode(study2, "Hamby173" = "Hamby44")) %>%
  mutate(across(is.factor, as.character)) %>%
  mutate(bullet1 = ifelse(barrel1 %in% letters, barrel1, bullet1),
         barrel1 = ifelse(barrel1 %in% letters, "Unknown", barrel1),
         bullet2 = ifelse(barrel2 %in% letters, barrel2, bullet2),
         barrel2 = ifelse(barrel2 %in% letters, "Unknown", barrel2),
         land1 = as.character(land1),
         land2 = as.character(land2),
         rfscore = predict(bulletxtrctr::rtrees, hamby173and252_raw %>%
                             select(rf_features), 
                           type = "prob")[,2]) %>%
  rename(samesource = match) %>%
  select(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2,
         rf_features, samesource, rfscore)
```

@@

```{r}
# Remove the comparisons involving the four lands with tank rash
hamby173and252_train <- hamby173and252_train_cleaning %>%
  mutate(bbl1 = paste(barrel1, bullet1, land1, sep = ":"),
         bbl2 = paste(barrel2, bullet2, land2, sep = ":")) %>%
  filter(!(bbl1 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4") | 
           bbl2 %in% c("6:2:1", "9:2:4", "Unknown:B:2", "Unknown:Q:4"))) %>%
  select(-bbl1, -bbl2)
```

@@

```{r}
# Create dataset for creating the parallel coordinate plot
par_data <- hamby173and252_train_cleaning %>%
  mutate(bbl1 = paste(barrel1, bullet1, land1, sep = ":"),
         bbl2 = paste(barrel2, bullet2, land2, sep = ":")) %>%
  mutate(category = factor(ifelse(!(bbl1 %in% c("6:2:1", "9:2:4", "Unknown:B:2",
                                                "Unknown:Q:4") | 
                                      bbl2 %in% c("6:2:1", "9:2:4", "Unknown:B:2",
                                                  "Unknown:Q:4")),
                           "keep", "remove"))) %>%
  arrange(category) %>%
  mutate(samesource_colors = ifelse(samesource == TRUE, rgb(1, 0, 0, alpha = 0.05), 
                                    rgb(0, 0, 0, alpha = 0.001)),
         samesource_colors2 = ifelse(samesource == TRUE, rgb(1, 0, 0, alpha = 0.001), 
                                    rgb(0, 0, 0, alpha = 0.01)),
         tankrash_colors = ifelse(category == "keep", rgb(0, 0, 1, alpha = 0.05),
                                  rgb(0, 0, 0, alpha = 0.01)))
```

The first plot shows the observations that were kept in black and the observations that were removed in blue. These observations do not stand out to me. It is not clear if these were the ones that were suppose to be removed or not. The second plot shows the observations from matches plotted in red, and the third plot shows the observations from non-matches plotted in black.

@@

Below is a parallel coordinate plot of the training dataset. It is faceted by `samesource` and colored by `rfscore`. The features are sorted by the variable importance scores (descending from left to right) from the `rtrees` model. It is interesting to see that the matches with low rfscores have feature values that are similar to the nonmatches.

@@

Below is a parallel coordinate plot of the testing dataset. It is faceted by `samesource` and colored by `rfscore`. You can see a clear distinction between the trends of the feature values from the nonmatches and matches. It is interesting to consider the cases in set 1 that are matches but were predicted to be nonmatches by the model. These are the cases colored in gray, and their feature values are very different from the observations that were predicted to be matches by the model. The nonmatch predictions have values that are closer to the nonmatches.

@@

I have been thinking a lot about it is difficult to determine whether LIME is providing reasonable explanations, because we cannot know what the random forest is doing. That is the point of LIME. Heike suggested that we fit a logistic regression model to the training bullet data. We can then apply LIME to the logistic regression and determine if the explanations match what we would expect based on the coefficients of the logistic regression that we can interpret. She also suggested including interactions in the logistic regression model. This section includes code for fitting several logistic regression models.

### 2020-09-25 journals/10-comparing-rfs/10-comparing-rfs.Rmd
everything is a mess right now, but pushing this so it doesn't get lost

# plot votes in a scatterplot
plot(rtrees2$votes, rtrees3$votes)
plot(rtrees2$votes[,2], bulletxtrctr::rtrees$votes[,2])
plot(rtrees3$votes[,2], bulletxtrctr::rtrees$votes[,2])

plot(sort(rtrees2$votes[,2]), sort(bulletxtrctr::rtrees$votes[,2]))
plot(sort(rtrees3$votes[,2]), sort(bulletxtrctr::rtrees$votes[,2]))
```

>>>>

# plot votes in a scatterplot
plot(rtrees2$votes, rtrees3$votes)
plot(rtrees2$votes, bulletxtrctr::rtrees$votes)
plot(rtrees3$votes, bulletxtrctr::rtrees$votes)

plot(sort(rtrees2$votes[,2]), sort(bulletxtrctr::rtrees$votes[,2]))
plot(sort(rtrees3$votes[,2]), sort(bulletxtrctr::rtrees$votes[,2]))

bulletxtrctr::rtrees$confusion
rtrees2$confusion
rtrees3$confusion
```

### 2020-09-25 journals/11-training-rtrees-with-CCFs_withlands/11-training-rtrees-with-CCFs_withlands.Rmd
everything is a mess right now, but pushing this so it doesn't get lost

---
title: "Identifying `rtrees` Training Data"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE, 
  dpi = 300
)
```

This journal contains work to identify the data used to train `rtrees` (contained in both `bulletr` and `bulletxtrctr`). We recently discovered some discrepancies between `rtrees` and the training data we were using (`hamby-comparisons.csv`, previously `features-hamby173and252.csv`). Unfortunately, the original work by Eric is not well documented, so it is difficult to know for sure which dataset was used to train the model. However, Eric recently pointed us to some [code](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R), which he believes was used to train `rtrees`. Heike was able to rerun part of the code to extract data from the CSAFE database that was created by the first part of the code (scans to features). The new data (`CCFs_withlands`) contains the same number of rows the number of predictions associated with `rtrees`. I have gathered all of our work trying to determine that `CCFs_withlands` is the data used to train `rtrees` in this journal.

@@

# Raw Datasets

This section contains code for loading and performing initial comparisons of the raw datasets.

## Load Data

Load the data provided by Heike and CSAFE that we have been using and believed to be the training data for rtrees:

```{r}
hamby_comparisons <- read.csv("../../../data/raw/hamby-comparisons.csv")
```

Load the data extracted (on 2020-09-23) from the CSAFE database and corresponds to the [R script](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R) Eric believes to be the one used to train `rtrees`.

```{r}
CCFs_withlands <- read.csv("../../../data/raw/CCFs_withlands.csv")
```

Obtain the features used to train `rtrees` for use throughout this journal:

```{r}
rtrees_features <- rownames(bulletxtrctr::rtrees$importance)
```

## Initial Comparisons

Dimensions of the raw datasets:

```{r}
dim(hamby_comparisons)
dim(CCFs_withlands)
```

Note that `CCFs_withlands` has the same number of rows as the number of predictions associated with `rtrees`:

```{r}
length(bulletxtrctr::rtrees$predicted)
```

Comparing summaries of the distributions of the features used to train `rtrees` from the two datasets. Note that I even remove the observations in `hamby_comparisons` known to have tank rash (`flag == FALSE`). The main differences are seen with the distributions of `D` (distance) and `sd_D` (standard deviation of distance).

```{r}
summary(hamby_comparisons %>% filter(flag == FALSE) %>% select(all_of(rtrees_features)))
summary(CCFs_withlands %>% select(all_of(rtrees_features)))
```

Comparing visualizations of the distributions of the features used to train `rtrees` from the two datasets to again see the differences in distributions of `D` and `sd_D`:

```{r fig.width = 10, fig.height = 3}
bind_rows(
  hamby_comparisons %>%
    filter(flag == FALSE) %>%
    select(all_of(rtrees_features)) %>%
    pivot_longer(cols = everything()) %>%
    mutate(dataset = "hamby_comparisons"),
  CCFs_withlands %>% select(all_of(rtrees_features)) %>%
    pivot_longer(cols = everything()) %>%
    mutate(dataset = "CCFs_withlands")
) %>%
  ggplot(aes(x = value, fill = dataset)) + 
  geom_histogram() +
  facet_grid(dataset ~ name)
```

<!-- New data plot: -->

<!-- ```{r} -->
<!-- preds <- -->
<!--   data.frame( -->
<!--     oob = predict(bulletxtrctr::rtrees, type = "prob")[, 2], -->
<!--     all_data = predict( -->
<!--       bulletxtrctr::rtrees, -->
<!--       ccfs_withlands_raw %>% select(all_of(rf_features)), -->
<!--       type = "prob" -->
<!--     )[, 2] -->
<!--   ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preds %>%  -->
<!--   ggplot(aes(x = all_data, y = oob)) +  -->
<!--   geom_point() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preds %>% -->
<!--   mutate(obs = 1:n()) %>% -->
<!--   pivot_longer(cols = -obs) %>% -->
<!--   ggplot(aes(x = value, fill = name)) +  -->
<!--   geom_density(alpha = 0.5) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preds %>%  -->
<!--   mutate(diff = oob - all_data) %>% -->
<!--   ggplot(aes(x = diff)) +  -->
<!--   geom_histogram() -->
<!-- ``` -->
<!-- # Data -->

<!-- Load the prepared training data which only has observations with `flag == FALSE` (prepared in the LIME diagnostics paper repository): -->

<!-- ```{r} -->
<!-- bullet_train <- read.csv("../../../diagnostics-paper/data/bullet-train.csv") -->
<!-- ``` -->

<!-- Load the random forest `rtrees` from the `bulletxtrctr` package and the random forest model I recently fit using the `bullet_train` data in the paper: -->

<!-- ```{r} -->
<!-- rtrees <- bulletxtrctr::rtrees -->
<!-- paper_rf <- readRDS("../../../diagnostics-paper/data/bullet-rf.rds") -->
<!-- ``` -->

<!-- Extract the features used to fit `rtrees`: -->

<!-- ```{r} -->
<!-- bullet_features <- rownames(bulletxtrctr::rtrees$importance) -->
<!-- bullet_features -->
<!-- ``` -->

<!-- # `rtrees` versus paper model -->

<!-- The code below was used to fit the random forest in the paper but is not run here: -->

<!-- ```{r eval = FALSE} -->
<!-- set.seed(20160104) -->
<!-- bullet_rf <- -->
<!--   randomForest( -->
<!--     y = factor(bullet_train$samesource), -->
<!--     x = bullet_train %>% select(all_of(bullet_features)), -->
<!--     ntree = bulletxtrctr::rtrees$ntree, -->
<!--     mtry = bulletxtrctr::rtrees$mtry -->
<!--   ) -->
<!-- ``` -->

<!-- Confusion matrices from the two models: -->

<!-- ```{r} -->
<!-- rtrees$confusion -->
<!-- paper_rf$confusion -->
<!-- ``` -->

<!-- Variable importance values: -->

<!-- ```{r} -->
<!-- vi = data.frame( -->
<!--   features = rownames(rtrees$importance), -->
<!--   rtrees = rtrees$importance[, 1], -->
<!--   paper = paper_rf$importance[, 1] -->
<!-- ) -->
<!-- ``` -->

<!-- # New Random Forests -->

<!-- Randomly select 100 seeds (using a random seed):  -->

<!-- ```{r} -->
<!-- set.seed(20200917) -->
<!-- seeds = sample(x = 10000000:100000000, size = 100, replace = FALSE) -->
<!-- seeds -->
<!-- ``` -->

<!-- Function for fitting random forests to mimic rtrees with different seeds: -->

<!-- ```{r} -->
<!-- fit_rf <- function(seed) { -->
<!--   set.seed(seed) -->
<!--   bullet_rf <- -->
<!--     randomForest( -->
<!--       y = factor(bullet_train$samesource), -->
<!--       x = bullet_train %>% select(all_of(bullet_features)), -->
<!--       ntree = bulletxtrctr::rtrees$ntree, -->
<!--       mtry = bulletxtrctr::rtrees$mtry -->
<!--     )   -->
<!-- } -->
<!-- ``` -->

<!-- Fit 100 new random forests and save or load: -->

<!-- ```{r} -->
<!-- # Fit (and save) or load random forests -->
<!-- bullet_rfs_file = "../../../data/bullet_rfs.rds" -->
<!-- if (!file.exists(bullet_rfs_file)) { -->
<!--   bullet_rfs = map(.x = seeds, .f = fit_rf) -->
<!--   saveRDS(object = bullet_rfs, file = bullet_rfs_file) -->
<!-- } else { -->
<!--   bullet_rfs = readRDS(bullet_rfs_file) -->
<!-- } -->
<!-- ``` -->

<!-- Function for extracting the confusion matrices from the model and put in a nice data frame: -->

<!-- ```{r} -->
<!-- extract_confus <- function(model) { -->
<!--   data.frame(model$confusion) %>% -->
<!--   rename("FALSE" = "FALSE.", "TRUE" = "TRUE.", "class_error" = "class.error") %>% -->
<!--   mutate(observed = rownames(model$confusion)) %>% -->
<!--   pivot_wider(names_from = observed, values_from = everything())   -->
<!-- } -->
<!-- ``` -->

<!-- Function for extracting the importance values, computing the rank of the features based on importance, and putting in a data frame:  -->

<!-- ```{r} -->
<!-- extract_importance <- function(model) { -->
<!--   data.frame(model$importance) %>% -->
<!--     mutate(features = rownames(model$importance)) %>% -->
<!--     arrange(desc(MeanDecreaseGini)) %>% -->
<!--     mutate(rank = 1:n()) %>% -->
<!--     rename("importance" = "MeanDecreaseGini") %>% -->
<!--     select(features, importance, rank) -->
<!-- } -->
<!-- ``` -->

<!-- Extract the confusion matrices and importance values from all random forest models: -->

<!-- ```{r} -->
<!-- res_confus <- map_df(.x = bullet_rfs, .f = extract_confus, .id = "rf_model_id") -->
<!-- res_importance <- map_df(.x = bullet_rfs, .f = extract_importance, .id = "rf_model_id") -->
<!-- ``` -->

<!-- # Visualizations of Results -->

<!-- ## Class Errors -->

<!-- Obtain the class errors from `rtrees` and the paper random forest: -->

<!-- ```{r} -->
<!-- comparison_values <- -->
<!--   data.frame( -->
<!--     model = c("paper", "rtrees"), -->
<!--     class_error_true = c(paper_rf$confusion[2, 3], rtrees$confusion[2, 3]), -->
<!--     class_error_false = c(paper_rf$confusion[1, 3], rtrees$confusion[1, 3]) -->
<!--   ) -->
<!-- ``` -->

<!-- Histograms of the class errors (true and false) show that the rtrees values appear to be in the extreme (false) or outside of the new random forest models (true): -->

<!-- ```{r fig.width = 10, fig.height = 4.5} -->
<!-- hist_true <-  -->
<!--   res_confus %>% -->
<!--   ggplot(aes(x = class_error_TRUE)) + -->
<!--   geom_histogram() + -->
<!--   geom_vline(data = comparison_values, -->
<!--              mapping = aes(xintercept = class_error_true, color = model)) -->

<!-- hist_false <-  -->
<!--   res_confus %>% -->
<!--   ggplot(aes(x = class_error_FALSE)) + -->
<!--   geom_histogram() + -->
<!--   geom_vline(data = comparison_values, -->
<!--              mapping = aes(xintercept = class_error_false, color = model)) -->

<!-- plot_grid(hist_true, hist_false) -->
<!-- ``` -->

<!-- Class error for TRUE versus class error for FALSE and again the rtrees observations falls outside the range of the newly trained random forest models: -->

<!-- ```{r} -->
<!-- ggplot() + -->
<!--   geom_point(data = res_confus, -->
<!--              mapping = aes(x = class_error_FALSE, y = class_error_TRUE)) + -->
<!--   geom_point( -->
<!--     data = comparison_values, -->
<!--     mapping = aes(x = class_error_false, y = class_error_true, color = model) -->
<!--   ) -->
<!-- ``` -->

>>>>

---
title: "Experiments with rtrees using `CCFs_withlands` data"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE
)
```

This journal contains the results from 100 new random forests to the bullet training data observations where `flag == FALSE`.

@@

# Data

Load the data Heike extracted from the CSAFE database (on 9/23) which we currently believe is the correct training data for `rtrees`:

```{r}
CCFs_withlands <- read.csv("../../../diagnostics-paper/data/raw/CCFs_withlands.csv")
```

Load the random forest `rtrees` from the `bulletxtrctr` package:

```{r}
rtrees <- bulletxtrctr::rtrees
```

Extract the features used to fit `rtrees`:

```{r}
bullet_features <- rownames(bulletxtrctr::rtrees$importance)
bullet_features
```

# Random Forests

Randomly select a large amount of seeds (using a random seed):

```{r}
set.seed(20200925)
seeds = sample(x = 10000000:100000000, size = 250, replace = FALSE)
seeds
```

Function for fitting random forests to mimic rtrees with different seeds:

```{r}
fit_rf <- function(seed) {
  set.seed(seed)
  bullet_rf <-
    randomForest(
      y = factor(CCFs_withlands$match),
      x = CCFs_withlands %>% select(all_of(bullet_features)),
      ntree = bulletxtrctr::rtrees$ntree,
      mtry = bulletxtrctr::rtrees$mtry
    )  
}
```

Fit 100 new random forests and save or load:

```{r}
# Fit (and save) or load random forests
CCFs_withlands_rfs_file = "../../../data/CCFs_withlands_rfs.rds"
if (!file.exists(CCFs_withlands_rfs_file)) {
  CCFs_withlands_rfs = map(.x = seeds, .f = fit_rf)
  saveRDS(object = CCFs_withlands_rfs, file = CCFs_withlands_rfs_file)
} else {
  CCFs_withlands_rfs = readRDS(CCFs_withlands_rfs_file)
}
```

Function for extracting the confusion matrices from the model and put in a nice data frame:

```{r}
extract_confus <- function(model) {
  data.frame(model$confusion) %>%
  rename("FALSE" = "FALSE.", "TRUE" = "TRUE.", "class_error" = "class.error") %>%
  mutate(observed = rownames(model$confusion)) %>%
  pivot_wider(names_from = observed, values_from = everything())  
}
```

Function for extracting the importance values, computing the rank of the features based on importance, and putting in a data frame:

```{r}
extract_importance <- function(model) {
  data.frame(model$importance) %>%
    mutate(features = rownames(model$importance)) %>%
    arrange(desc(MeanDecreaseGini)) %>%
    mutate(rank = 1:n()) %>%
    rename("importance" = "MeanDecreaseGini") %>%
    select(features, importance, rank)
}
```

Extract the confusion matrices and importance values from all random forest models:

```{r}
res_confus <- map_df(.x = CCFs_withlands_rfs, .f = extract_confus, .id = "rf_model_id")
res_importance <- map_df(.x = CCFs_withlands_rfs, .f = extract_importance, .id = "rf_model_id")
```

# Visualizations of Results

## Class Errors

Obtain the class errors from `rtrees` and the paper random forest:

```{r}
comparison_values <-
  data.frame(
    model = "rtrees",
    class_error_true = rtrees$confusion[2, 3],
    class_error_false = rtrees$confusion[1, 3]
  )
```

Histograms of the class errors (true and false) show that the rtrees values appear to be in the extreme (false) or outside of the new random forest models (true):

```{r fig.width = 10, fig.height = 4.5}
hist_true <- 
  res_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram() +
  geom_vline(data = comparison_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false <- 
  res_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram() +
  geom_vline(data = comparison_values,
             mapping = aes(xintercept = class_error_false, color = model))

plot_grid(hist_true, hist_false)
```

Class error for TRUE versus class error for FALSE and again the rtrees observations falls outside the range of the newly trained random forest models:

```{r}
ggplot() +
  geom_point(data = res_confus,
             mapping = aes(x = class_error_FALSE, y = class_error_TRUE)) +
  geom_point(
    data = comparison_values,
    mapping = aes(x = class_error_false, y = class_error_true, color = model)
  )
```
